# Tweedie

*This file contains illustrative **R** code for the Tweedie distribution. When reviewing this code, you should open an **R** session, copy-and-paste the code, and see it perform. Then, you will be able to change parameters, look up commands, and so forth, as you go. *

## Tweedie distribution{.tabset}

### Load Tweedie Package
First bring in the package Tweedie (you may need to first install this package).
```{r warning=FALSE , message=FALSE , comment=""}
library(tweedie)
```

### Set Paramteres for Tweedie(p,mu,phi)
Setting parameters p, mu and phi defines the specific features of the distribution.  

Furthermore, setting a specific seed allows us to generate the same randomn numbers so we can produce identical distributions
```{r warning=FALSE , message=FALSE , comment=""}
set.seed(123)
p <- 1.5
mu <- exp(1) 
phi <- exp(1)
```

### Set Sample Size
Sample size is set to 500 for this example.
"y" holds all 500 obserations from tweedie distribution with the given parameters.
```{r warning=FALSE , message=FALSE , comment=""}
n <- 500
y <- rtweedie(n,p,mu,phi)
```

### Show Summary Statistics
Here we calculate important statisitics like mean, median, standard deviation and quantiles.
```{r warning=FALSE , message=FALSE , comment=""}
summary(y)
sd(y)
quantile(y,seq(0,1,0.1))
```

### Show Histogram
Histograms are useful for visually interpreting data.
Sometime summary statistics aren't enough to see the full picture.
```{r warning=FALSE , message=FALSE , comment=""}
hist(y, prob=T,breaks=50)
x <- seq(0,max(y),0.1)
lines(x,dtweedie(x,p,mu,phi),col="red")
```

### QQ Plots for Different p Values
A QQ plot is a plot of the quantiles of the first data set against the quantiles of the second data set.

This is graphical technique for determining if two data sets come from populations with a common distribution.

It appears here that a power of 1.5 matches the distribution best.
```{r warning=FALSE , message=FALSE , comment=""}
par(mfrow=c(2,2),mar=c(4,4,4,4))
qqTweedie <- function(xi,pct,mu,phi) {
  plot(qtweedie(pct,xi,mu,phi),quantile(y,probs=pct),
       main=paste("Power = ",xi), xlab="Theoretical Quantile", ylab="Empirical Quantile")
  abline(0,1)
}
pct <- seq(0.01,0.99,0.01)
lapply(c(1,1.5,2,2.5),qqTweedie,pct=pct,mu=mu,phi=phi)
```


### Fit Tweedie Distribution
Here we run a  <a href="#" data-toggle="tooltip" title="Generalized Linear Model", style="color:green;">"glm"</a> for the Tweedie distribution. 
you may need to first install the "statmod" package
```{r warning=FALSE , message=FALSE , comment=""}
library(statmod)
fit <- glm(y~1,family=tweedie(var.power=1.5,link.power=0))
summary(fit)
```

### Show Parameter Estimates
We now display the parameter estimates calculated in the glm.
```{r warning=FALSE , message=FALSE , comment=""}
summary(fit)$coefficient
summary(fit)$dispersion
```

### Maximum Likelihood Estimation
Here we run a <a href="#" data-toggle="tooltip" title="Maximum Likelihood Estimate", style="color:green;">"MLE"</a> to determine the most likely parameters of the Tweedie distribution.
```{r warning=FALSE , message=FALSE , comment=""}
loglik<-function(parms){ 
  p=parms[1]
  mu=exp(parms[2])
  phi=exp(parms[3])
  llk <- -sum(log(dtweedie(y, p, mu, phi)))
  llk
}
ini <- c(1.5,1,1)
zop <- nlminb(ini,loglik, lower =c(1+1e-6,-Inf,-Inf),upper =c(2-1e-6,Inf,Inf))
print(zop)
```

### Obtain Standard Error
Now we calculate the standard errors of our parameter estimates from the MLE.
You may need to first install the "numDeriv" package.
```{r warning=FALSE , message=FALSE , comment=""}
library(numDeriv)
est <- zop$par
names(est) <- c("p","mu","phi")
hess<-hessian(loglik,est)
se <-sqrt(diag(solve(hess)))
print(cbind(est,se))
```

