# Bootstrap Estimation

*This file demonstrates both empirical and parametric boostrap simulation. When reviewing this code, you should open an **R** session, copy-and-paste the code, and see it perform. Then, you will be able to change parameters, look up commands, and so forth, as you go.*

## Empirical Bootstrap {.tabset}

Example: 90% confidence interval for the mean. Consider outcomes of rolling a fair die.

### Random sample 

Here we input a sample of 10 observations, which we are going to build our estimates off of.
```{r comment=""}
y <- c(1,3,2,5,4,5,5,6,6,6)
n <- length(y)
n
```

### Sample mean 

Finding the mean of the random sample.
```{r comment=""}
xbar <- mean(y)
xbar
```

### Random resamples from y
#### Set bootstrap sample size

We're going to generate 30 different observations using the original sample data, called the *boostrap sample*
```{r comment=""}
nboot <- 30
tmpdata <- sample(y,n*nboot,replace=TRUE)
bootstrap.sample <- matrix(tmpdata,nrow=n,ncol=nboot)
```

#### Compute sample mean for each bootstrap sample

Here we find the mean for all 30 of our boostrap samples.
```{r comment=""}
bsmeans <- colMeans(bootstrap.sample)
bsmeans
```

### 90% confidence interval

Using the generated boostrap sample, we calculate a 90% confidence interval for our sample mean.
```{r comment=""}
CI <- c(quantile(bsmeans,prob=0.05), quantile(bsmeans,prob=0.95))
CI
```

## Parametric Bootstrap {.tabset}

Example: confidence interval for 1/theta. Consider y ~ exponential(theta=10).

### random sample of size 250

This time we generate a random sample of 250 observations from the exponential sampling distribution.
```{r comment=""}
y <- rexp(250,rate=0.1)
n <- length(y)
n
```

### The MLE for lambda: 1/xbar

Remember from earlier that the MLE of theta for an exponential distribution is its mean.
```{r comment=""}
theta.mle <- mean(y)
rate.hat <- 1/theta.mle
```

### Generate bootstrap samples

Using the MLE we calculated we now generate 500 *bootstrap samples* of 250 observations each.
```{r comment=""}
nboot <- 500
tmpdata <- rexp(n*nboot,rate=rate.hat)
bootstrap.sample <- matrix(tmpdata,nrow=n,ncol=nboot)
```

### compute botstrap statistics

We now find the rate parameter for each of our *boostrap samples*
```{r comment=""}
rate.star <- 1/colMeans(bootstrap.sample)
```

### calculate deviation from sample statistics

Subtracting the original sample rate parameter from each of our simulated rate parameters, we can find the 5th and 95th percentiles of the differences.
```{r comment=""}
delta.star <- rate.star - rate.hat
delta.lb <- quantile(delta.star,prob=0.05)
delta.ub <- quantile(delta.star,prob=0.95)
```

### 90% confidence intervel

We use the percentiles to create a 90% CI for the rate parameter.
```{r comment=""}
CI <- rate.hat - c(delta.ub,delta.lb)
print(CI)
```  
