---
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=4"
    ]
---

# Simulation 

*This file contains illustrative **R** code for computing important count distributions. When reviewing this code, you should open an **R** session, copy-and-paste the code, and see it perform. Then, you will be able to change parameters, look up commands, and so forth, as you go. *

## Simulation - Inversion Method
This section shows how to use the inversion method to simulate claims from a gamma distribution. The results below are summary statistics from the simulated data. 

```{r}
##Simulation-gamma
library(moments)
set.seed(2) #set seed to reproduce work 
nTot<-20000  #number of simulations
alpha<-2
theta<-100
         
Losses<-rgamma(nTot,alpha,scale = theta)  
summary(Losses)

k<-0.95
PercentileLoss<-quantile(Losses,k)     # Kth percentile of Losses 
PercentileLoss

#######################################
#OR you can use this method to simulate losses 
#Fx<-runif(nTot)
#Losses<-qgamma(Fx,alpha,scale=theta)
#######################################
# For the Pareto Distribution, use
#library(VGAM)
# nTot<-10000  #number of simulations
# alpha<-3
# theta<-25000
# Losses<-rparetoII(nTot,scale=theta,shape=alpha)
# rparetoII(nTot,scale=theta,shape=alpha) 
```
A few quick notes on these commands:

* The `rgamma()` function randomly generates data from the Gamma distribution.  In this illustration the data was generated from a gamma distribution with parameters `shape=alpha=2` and `scale=theta=100`. 
* The `quantile()` function provides sample quantiles corresponding to the given probabilities. Here we wanted the simulated loss data corresponding to the 95th percentile. 



## Comparing Moments from The Simulated Data to Theoretical Moments

```{r}
library(pander)
###Raw moments for k=0.5
# Theoretical 
k<-0.5
T0.5<-round(((theta^k)*gamma(alpha+k))/gamma(alpha),2)

#Simulated data raw moments
S0.5<-round(moment(Losses, order = k, central = FALSE),2)


###Raw moments for k=1
# Theoretical 
k<-1
T1<-((theta^k)*gamma(alpha+k))/gamma(alpha)

#Simulated data raw moments
S1<-round(moment(Losses, order = k, central = FALSE),2)

###Raw moments for k=2
# Theoretical 
k<-2
T2<-((theta^k)*gamma(alpha+k))/gamma(alpha)

#Simulated data raw moments
S2<-round(moment(Losses, order = k, central = FALSE),2)

###Raw moments for k=3
# Theoretical 
k<-3
T3<-((theta^k)*gamma(alpha+k))/gamma(alpha)

#Simulated data raw moments
S3<-round(moment(Losses, order = k, central = FALSE),2)

###Raw moments for k=3
# Theoretical 
k<-4
T4<-((theta^k)*gamma(alpha+k))/gamma(alpha)

#Simulated data raw moments
S4<-round(moment(Losses, order = k, central = FALSE),2)

pander(rbind(c("k",0.5,1,2,3,4),c("Theoritical",T0.5,T1,T2,T3,T4),c("Simulated",S0.5,S1,S2,S3,S4)))
```
<br/>
A few quick notes on these commands:

* The `momemt()` function computes all the sample moments of the chosen type up to a given order. In this illustration, we wanted the raw sample moments of the *kth* order.
* The `round()` function was used to round values. 

