# Aggregate Claim Simulation

*This file demonstrates simulation of aggregate claim distributions. When reviewing this code, you should open an **R** session, copy-and-paste the code, and see it perform. Then, you will be able to change parameters, look up commands, and so forth, as you go.*

## Collective Risk Model: without coverage modifications {.tabset}

S = X_1 + ... + X_N

Assume N ~ Poisson(lambda=2) and X ~ Pareto(alpha=3,theta=5000)

### Set Parameters
```{r comment=""}
lambda <- 2
alpha <- 3
theta <- 5000
```

### Show frequency and severity distributions
Graphing the our frequency (N) and severity (X) distributions

```{r comment=""}
par(mfrow=c(1,2))

n <- 1:10
fn <- dpois(1:10,lambda)

plot(n,fn,ylim=c(0,0.3),main="Frequency: Poisson")
abline(h=0,lty=2)

x <- seq(1,25000,1)
fx <- alpha*theta^alpha/(x+theta)^(alpha+1)

plot(x,fx,type="l",main="Severity: Pareto")
```

### Set sample size for the simulation
We're going to simulate 5000 observations of S

```{r comment=""}
set.seed(123)
size <- 5000
S <- rep(NA,size)
N <- rpois(size,lambda)
for (i in 1:size){
  uu <- runif(N[i])
  X <- theta*((1-uu)^(-1/alpha)-1)
  S[i] <- sum(X)
}
```

### Show distribution of aggregate loss S

```{r comment=""}
par(mfrow=c(1,2))
hist(S,freq=F,breaks=100)
plot(ecdf(S),xlab="S")
```

## Applications {.tabset}

### Find descriptive statistics
Here we show numerical descriptions of our simulated distribution S

```{r comment=""}
mean(S)                             # sample mean
sd(S)                               # sample standard deviation
quantile(S,prob=c(0.05,0.5,0.95))   # percentiles
```

### Calculate cdf

```{r comment=""}
sum((S==0))/size
```
Pr(S=0)

```{r comment=""}
sum(S<=mean(S))/size
```
Pr(S<=E(S))

```{r comment=""}
sum(S>mean(S))/size 
```
Pr(S>E(S))

### Calculate risk measures
<a href="#" data-toggle="tooltip" title="Conditional Tail Expectation", style="color:green;">CTE</a> is also known as <a href="#" data-toggle="tooltip" title="Tail Value at Risk", style="color:green;">TVaR</a>

<script>
$(document).ready(function(){
    $('[data-toggle="tooltip"]').tooltip();
});
</script>

```{r comment=""}
VaR <- quantile(S,prob=0.99)         # significance level = 0.01
CTE <- sum(S*(S>VaR))/sum((S>VaR))
rm <- c(VaR,CTE)
names(rm) <- c("VaR","CTE")
print(rm)
```

### Pricing stop-loss insurance - Set deductible
Here we plot how the premium for a stop-loss insurance product changes based on the size of the deductible

```{r comment=""}
par(mfrow=c(1,1))
d <- seq(0,120000,1000)
price <- rep(NA,length(d))
for (i in 1:length(d)){
  price[i] = sum((S-d[i])*(S>d[i]))/size
}
plot(d,price,xlab="Deductible",ylab="Stop-Loss Premium",type="b")
```