---
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=2"
    ]
---

# Modeling Loss Severities


*This file contains illustrative **R** code for computing important count distributions. When reviewing this code, you should open an **R** session, copy-and-paste the code, and see it perform. Then, you will be able to change parameters, look up commands, and so forth, as you go. *

## Required packages
```{r warning=FALSE, message=FALSE}
library(actuar)
library(VGAM)
```

## Gamma Distribution {.tabset}
This section demonstrates the effect of the shape and scale parameters on the gamma density. 

### Varying The Shape Parameter
The graph shows the gamma density functions with varying shape parameters $(\alpha)$

```{r}
# Example 1: gamma distribution
# define a grid
x <- seq(0,1000,by=1)

# define a set of scale and shape parameters
scaleparam <- seq(100,250,by=50)
shapeparam <- 2:5

# varying the shape parameter
plot(x, dgamma(x, shape = shapeparam[1], scale = 100), type = "l", ylab = "Gamma density")

for(k in 2:length(shapeparam)){
  lines(x,dgamma(x,shape = shapeparam[k], scale = 100), col = k)
}
legend("topright", c(expression(alpha~'=2'), expression(alpha~'=3'), expression(alpha~'=4'), expression(alpha~'=5')), lty=1, col = 1:4)
title(substitute(paste("Pdf Gamma Density with"," ",theta,"=100"," ", "and Varying Shape")))
```

A few quick notes on these commands :

* `dgamma` function is used for density of the gamma distribution with shape and scale parameters . 
* The `legend` function can be used to add legends to plots. 

* <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-placement="top" data-trigger="hover" data-content="A very handy command for displaying results graphically.", style="color:green;">`plot`</a>
 <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-trigger="hover" data-content="Is very handy for superimposing one graph on another.", style="color:green;">`lines`</a>


### Varying The Scale Parameter
The graph shows the gamma density functions with varying scale parameters $(\theta)$

```{r}
plot(x, dgamma(x, shape = 2, scale = scaleparam[1]), type = "l", ylab = "Gamma density")

for(k in 2:length(scaleparam)){
  lines(x,dgamma(x,shape = 2, scale = scaleparam[k]), col = k)
}

legend("topright", c(expression(theta~'=100'), expression(theta~'=150'), expression(theta~'=200'), expression(theta~'=250')), lty=1, col = 1:4)
title(substitute(paste("Pdf Gamma Density with"," ",alpha,"=2"," ", "and Varying Scale")))
```

```{r}
knitr::include_app("https://luyang.shinyapps.io/gamma/", 
  height = "600px")
```


## Pareto Distribution {.tabset}
This section demonstrates the effect of the shape and scale parameters on the Pareto density function.

### Varying The Shape Parameter
The graph shows the Pareto density functions with varying shape parameters $(\alpha)$
```{r}
z<- seq(0,3000,by=1)

scaleparam <- seq(2000,3500,500)
shapeparam <- 1:4

# varying the shape parameter
plot(z, dparetoII(z, loc=0, shape = shapeparam[1], scale = 2000), ylim=c(0,0.002),type = "l", ylab = "Pareto density")

for(k in 2:length(shapeparam)){
  lines(z,dparetoII(z,loc=0, shape = shapeparam[k], scale = 2000), col = k)
}
legend("topright", c(expression(alpha~'=1'), expression(alpha~'=2'), expression(alpha~'=3'), expression(alpha~'=4')), lty=1, col = 1:4)
title(substitute(paste("Pdf Pareto Density with"," ",theta,"=2000"," ", "and Varying Shape")))
```

### Varying The Scale Parameter
The graph shows the Pareto density functions with varying scale parameters $(\theta)$

```{r}
plot(z, dparetoII(z, loc=0, shape = 3, scale = scaleparam[1]), type = "l", ylab = "Pareto density")

for(k in 2:length(scaleparam)){
  lines(z,dparetoII(z,loc=0, shape = 3, scale = scaleparam[k]), col = k)
}

legend("topright", c(expression(theta~'=2000'), expression(theta~'=2500'), expression(theta~'=3000'), expression(theta~'=3500')), lty=1, col = 1:4)
title(substitute(paste("Pdf Pareto Density with"," ",alpha,"=3"," ", "and Varying Scale")))
```

## Weibull Distribution {.tabset}
This section demonstrates the effect of the shape and scale parameters on the Weibull density function.

### Varying The Shape Parameter
The graph shows the Weibull density function with varying shape parameters $(\alpha)$

```{r}
z<- seq(0,400,by=1)

scaleparam <- seq(50,200,50)
shapeparam <- seq(1.5,3,0.5)

# varying the shape parameter
plot(z, dweibull(z, shape = shapeparam[1], scale = 100), ylim=c(0,0.012), type = "l", ylab = "Weibull density")

for(k in 2:length(shapeparam)){
  lines(z,dweibull(z,shape = shapeparam[k], scale = 100), col = k)
}

legend("topright",  c(expression(alpha~'=1.5'), expression(alpha~'=2'), expression(alpha~'=2.5'), expression(alpha~'=3')), lty=1, col = 1:4)
title(substitute(paste("Pdf Weibull Density with"," ",theta,"=100"," ", "and Varying Shape")))

```

### Varying The Scale Parameter
The graph shows the Weibull density function with varying scale parameters $(\theta)$

```{r}
plot(z, dweibull(z, shape = 3, scale = scaleparam[1]), type = "l", ylab = "Weibull density")

for(k in 2:length(scaleparam)){
  lines(z,dweibull(z,shape = 3, scale = scaleparam[k]), col = k)
}
legend("topright", c(expression(theta~'=50'), expression(theta~'=100'), expression(theta~'=150'), expression(theta~'=200')), lty=1, col = 1:4)
title(substitute(paste("Pdf Weibull Density with"," ",alpha,"=3"," ", "and Varying Scale")))
```


## Generalized Beta Distribution of The Second Kind (GB2) {.tabset}
This section demonstrates the effect of the shape and scale parameters on the GB2 density function.

### Varying The Scale Parameter
The graph shows the GB2 density function with varying scale parameter $(\theta)$
```{r}
## Example 4:GB2
gb2density <- function(x,shape1,shape2,shape3,scale){
  mu <- log(scale)
  sigma <- 1/shape3
  xt <- (log(x)-mu)/sigma
  logexpxt<-ifelse(xt>23,yt,log(1+exp(xt)))
  logdens <- shape1*xt - log(sigma) - log(beta(shape1,shape2)) - (shape1+shape2)*logexpxt -log(x) 
  exp(logdens)
}
x<- seq(0,400,by=1)

alpha1<-5
alpha2<-4 
gamma <-2
theta <- seq(150,250,50)

# varying the scale parameter
plot(x, gb2density(x, shape1=alpha1,shape2=alpha2,shape3=gamma, scale = theta[1]), 
     type = "l", ylab = "Gen Beta 2 density",
     main = 
     expression(paste("GB2 Density with ", alpha[1], "=5,", alpha[2], "=4,", alpha[3], 
                      "=2, and Varying Scale (",theta, ") Parameters")) )

for(k in 2:length(theta)){
  lines(x,gb2density(x,shape1=alpha1,shape2=alpha2,shape3=gamma, scale = theta[k]), col = k)
}
legend("topleft", c(expression(theta~'=150'), expression(theta~'=200'), expression(theta~'=250')), lty=1, cex=0.6,col = 1:3)

```

Note: Here we wrote our own function for the density function of the GB2 density function. 

## Methods of Creating New Distributions {.tabset}
This section shows some of the methods of creating new distributions.

### Mixture Distributions
The graph below creates a density function from two random variables that follow a gamma distribution. 

```{r}
## Example 5: A mixed density
## specify density of a mixture of 2 gamma distributions
MixtureGammaDensity <- function(x, a1, a2, alphaGamma1, thetaGamma1, alphaGamma2, thetaGamma2){
  a1 * dgamma(x, shape = alphaGamma1, scale = thetaGamma1) + a2 * dgamma(x, shape = alphaGamma2, scale = thetaGamma2)
}

w <- 1:30000/100
a1<-0.5
a2<-0.5
alpha1 <- 4
theta1 <- 7
alpha2 <- 15
theta2 <- 7

MixGammadens <- MixtureGammaDensity(w, a1,a2,alpha1, theta1, alpha2, theta2)

plot(w, MixGammadens, type = "l")

```

### Density Obtained Through Splicing 
The graph below shows a density function through <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-trigger="hover" data-content="Join together different probability density functions to form a pdf over the support of a random variable.", style="color:green;">splicing</a> by combining an exponential distribution on $(0,c)$ with a Pareto distribution on $(c,\infty)$

```{r}
##Example 6: density obtained through splicing
## combine an Exp on (0,c) with a Pareto on (c,\infty)

SpliceExpPar <- function(x, c, v, theta, gamma, alpha){
  if(0<=x & x<c){return(v * dexp(x, 1/theta)/pexp(c,1/theta))}else
    if(x>=c){return((1-v)*dparetoII(x,loc=0, shape = alpha, scale = theta)/(1-pparetoII(x,loc=0, shape = alpha, scale = theta)))}
}

x <- t(as.matrix(1:2500/10))

spliceValues <- apply(x,2,SpliceExpPar, c = 100, v = 0.6, theta = 100, gamma = 200, alpha = 4)

plot(x,spliceValues, type = 'l')
```

## Coverage Modifications 

### Load Required Package 
 The `actuar` package provides functions for dealing with coverage modifications. In the following sections we will check the functionalities of the `coverage` command. 
 
```{r warning=FALSE , message=FALSE}
library(actuar)
```

### Ordinary Deductible {.tabset}
This section plots the modified probability density functions due to deductibles for the <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-trigger="hover" data-content="Zero is observed for losses less than the deductible.", style="color:green;">payment per loss</a> and <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-trigger="hover" data-content="If the loss is less than the deductible, then the insurer does not observe the loss.", style="color:green;">payment per payment</a> random variables.

#### Payment Per Loss with Ordinary Deductible
Let $X$ be the random variable for loss size. The random variable for the payment per loss with deductible $d$ is $Y^L=(X-d)_+$. The plot of the modified probability density function is below. 

```{r}
f <- coverage(dgamma, pgamma, deductible = 1, per.loss = TRUE)# create the object
mode(f) # it's a function. Here deductible is 1
### Check the pdf for Y^L at 0 and the original loss at 1
f(0, 3) # mass at 0
pgamma(0+1, 3) # idem

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3), lwd=1, col="gray") # original
curve(dgamma(x, 3), from = 1, to = 10, ylim = c(0, 0.3), lwd=2, add=TRUE)
curve(f(x, 3), from = 0.01, col = "blue", add = TRUE, lwd=2) # modified
points(0, f(0, 3), pch = 16, col = "blue")
legend("topright", c("Original pdf", "Modified pdf"), lty=1, cex=0.6,col = c("black","blue"))

```

A few quick notes on these commands:

* The `coverage()` function computes probability density function or cumulative distribution function of the payment per payment or payment per loss random variable under any combination of the following coverage modifications: deductible, limit, coinsurance, inflation. In this illustration we used it to compute the probability density function of the payment per loss random variable with a deductible of 1. 
* The `f(0, 3)` function calculates the pdf when the payment per loss variable is 0 with gamma parameters shape=3 and rate=1. Because we used a deductible of 1 , this should be equal to `pgamma(0+1, 3)`. 

#### Payment Per Payment with Ordinary Deductible

$Y^P$ with pdf $f_{Y^P}(y) = f_X(y+d)/S_X(d)$

```{r}

f <- coverage(dgamma, pgamma, deductible = 1) # create the object

f(0, 3) # calculate in x = 0, shape=3, rate=1
f(5, 3) # calculate in x = 5, shape=3, rate=1

dgamma(5 + 1, 3)/pgamma(1, 3, lower = FALSE) # DIY
curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3), lwd=1,col="gray")  # original pdf
curve(dgamma(x, 3), from = 1, to = 10, ylim = c(0, 0.3), add=TRUE, lwd=2) 
curve(f(x, 3), from = 0.01, col = "blue", add = TRUE,lwd=2)     # modified pdf
legend("topright", c("Original pdf", "Modified pdf"), lty=1, cex=0.6,col = c("black","blue"))

```

#### Per Payment Variable with Policy Limit, Coinsurance and Inflation 
```{r}
f <- coverage(dgamma, pgamma, deductible = 1, limit = 100, coinsurance = 0.9, inflation = 0.05) # create the object

f(0, 3) # calculate in x = 0, shape=3, rate=1
f(5, 3) # calculate in x = 5, shape=3, rate=1

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3), lwd=1,col="gray")# original pdf
curve(dgamma(x, 3), from = 1, to = 10, ylim = c(0, 0.3), add=TRUE, lwd=2)
curve(f(x, 3), from = 0.01, col = "blue", add = TRUE,lwd=2)     # modified pdf
legend("topright", c("Original pdf", "Modified pdf"), lty=1, cex=0.6,col = c("black","blue"))

``` 

### Franchise Deductible {.tabset}
A policy with a *franchise deductible* of $d$ pays nothing if the loss is no greater than $d$, and pays the full amount of the loss if it is greater than $d$. This section plots the pdf for the per payment and per loss random variable. 

#### Payment Per Loss with Franchise Deductible 

```{r}
# franchise deductible
# per loss variable
f <- coverage(dgamma, pgamma, deductible = 1,
per.loss = TRUE, franchise = TRUE)
f(0, 3) 	  # mass at 0
pgamma(1, 3)  # idem
f(0.5, 3) # 0 < x < 1
f(1, 3)   # x = 1
f(5, 3)   # x > 1
dgamma(5,3)

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3)) # original
curve(f(x, 3), from = 1.1, col = "blue", add = TRUE) # modified
points(0, f(0, 3), pch = 16, col = "blue") # mass at 0
curve(f(x, 3), from = 0.1, to = 1, col = "blue", add = TRUE) # 0 < x < 1
legend("topright", c("Original pdf", "Modified pdf"), lty=1, cex=0.6,col = c("black","blue"))
```

Note : to use the franchise deductible , we have to add the option `franchise = TRUE` in the coverage function. 

#### Payment Per Payment with Franchise Deductible
```{r}
# franchise deductible
# per payment variable
f <- coverage(dgamma, pgamma, deductible = 1, franchise = TRUE)
f(0, 3)   # x = 0
f(0.5, 3) # 0 < x < 1
f(1, 3) # x = 1
f(5, 3) # x > 1

dgamma(5, 3)/pgamma(1, 3, lower = FALSE) # idem

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3)) 	# original
curve(f(x, 3), from = 1.1, col = "blue", add = TRUE) 		# modified
curve(f(x, 3), from = 0, to = 1, col = "blue", add = TRUE)  # 0 < x < 1
legend("topright", c("Original pdf", "Modified pdf"), lty=1, cex=0.6,col = c("black","blue"))
```

<script>
$(document).ready(function(){
    $('[data-toggle="popover"]').popover(); 
});
</script>

