---
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=2"
    ]
---
# Modeling Loss Severities

*This file contains illustrative **R** code for computing important count distributions. When reviewing this code, you should open an **R** session, copy-and-paste the code, and see it perform. Then, you will be able to change parameters, look up commands, and so forth, as you go. *

## Required packages
```{r warning = FALSE, message = FALSE}
library(actuar)
library(VGAM)
```

## Gamma Distribution {.tabset}
This section demonstrates the effect of the shape and scale parameters on the gamma density. 

### Varying The Shape Parameter
The graph shows the gamma density functions with varying shape parameters $(\alpha)$

```{r}
# Example 1: gamma distribution
# Define a grid
x <- seq(0, 1000, by = 1)

# Define a set of scale and shape parameters
scale_param <- seq(100, 250, by = 50)
shape_param <- 2:5

# Varying the shape parameter
plot(x, dgamma(x, shape = shape_param[1], scale = 100), type = "l", ylab = "Gamma Density")

for (k in 2:length(shape_param)) {
  lines(x, dgamma(x, shape = shape_param[k], scale = 100), col = k)
}
legend("topright", c(expression(alpha ~ '= 2'), expression(alpha ~ '= 3'), 
                     expression(alpha ~ '= 4'), expression(alpha ~ '= 5')), 
       lty = 1, col = 1:4)
title(substitute(paste("Pdf Gamma Density with"," ",theta," = 100"," ", 
                       "and Varying Shape")))
```

A few quick notes on these commands :

* `seq` is short-hand for sequence
* `dgamma` function is used for density of the Gamma distribution with shape and scale parameters . 
* `plot` is a very handy command for displaying results graphically.
* The `lines()` function is used to add plots to an already existing graph. 
* The `legend` function can be used to add legends to plots.


### Varying The Scale Parameter
The graph shows the gamma density functions with varying scale parameters $(\theta)$

```{r}
plot(x, dgamma(x, shape = 2, scale = scale_param[1]), type = "l", ylab = "Gamma Density")

for (k in 2:length(scale_param)) {
  lines(x, dgamma(x, shape = 2, scale = scale_param[k]), col = k)
}

legend("topright", c(expression(theta ~ '= 100'), expression(theta ~ '= 150'), 
                     expression(theta ~'= 200'), expression(theta ~ '= 250')), 
       lty = 1, col = 1:4)
title(substitute(paste("Pdf Gamma Density with"," ", alpha, " = 2", " ", 
                       "and Varying Scale")))
```

```{r echo = FALSE}
knitr::include_app("https://luyang.shinyapps.io/gamma/", 
  height = "600px")
```


## Pareto Distribution {.tabset}
This section demonstrates the effect of the shape and scale parameters on the Pareto density function.

### Varying The Shape Parameter
The graph shows the Pareto density functions with varying shape parameters $(\alpha)$
```{r}
z <- seq(0, 3000, by = 1)

scale_param <- seq(2000, 3500, 500)
shape_param <- 1:4

# Varying the shape parameter
plot(z, dparetoII(z, loc = 0, shape = shape_param[1], scale = 2000), 
     ylim = c(0, 0.002),type = "l", ylab = "Pareto Density")

for (k in 2:length(shape_param)) {
  lines(z, dparetoII(z,loc = 0, shape = shape_param[k], scale = 2000), col = k)
}
legend("topright", c(expression(alpha ~ '= 1'), expression(alpha ~ '= 2'), 
                     expression(alpha ~ '= 3'), expression(alpha ~ '= 4')), 
       lty = 1, col = 1:4)
title(substitute(paste("Pdf Pareto Density with"," ",theta," = 2000"," ", 
                       "and Varying Shape")))
```

### Varying The Scale Parameter
The graph shows the Pareto density functions with varying scale parameters $(\theta)$

```{r}
plot(z, dparetoII(z, loc = 0, shape = 3, scale = scale_param[1]), 
     type = "l", ylab = "Pareto Density")

for (k in 2:length(scale_param)) {
  lines(z, dparetoII(z, loc = 0, shape = 3, scale = scale_param[k]), col = k)
}

legend("topright", c(expression(theta ~ '= 2000'), expression(theta ~ '= 2500'), 
                     expression(theta ~ '= 3000'), expression(theta ~ '= 3500')), 
       lty = 1, col = 1:4)
title(substitute(paste("Pdf Pareto Density with"," ",alpha," = 3"," ", 
                       "and Varying Scale")))
```

## Weibull Distribution {.tabset}
This section demonstrates the effect of the shape and scale parameters on the Weibull density function.

### Varying The Shape Parameter
The graph shows the Weibull density function with varying shape parameters $(\alpha)$

```{r}
z <- seq(0, 400, by = 1)

scale_param <- seq(50, 200, 50)
shape_param <- seq(1.5, 3, 0.5)

# Varying the shape parameter
plot(z, dweibull(z, shape = shape_param[1], scale = 100), 
     ylim = c(0, 0.012), type = "l", ylab = "Weibull Density")

for (k in 2:length(shape_param)) {
  lines(z, dweibull(z, shape = shape_param[k], scale = 100), col = k)
}

legend("topright",  c(expression(alpha ~ '= 1.5'), expression(alpha ~ '= 2'), 
                      expression(alpha ~'= 2.5'), expression(alpha ~ '= 3')), 
       lty = 1, col = 1:4)
title(substitute(paste("Pdf Weibull Density with"," ",theta," = 100"," ", 
                       "and Varying Shape")))

```

### Varying The Scale Parameter
The graph shows the Weibull density function with varying scale parameters $(\theta)$

```{r}
plot(z, dweibull(z, shape = 3, scale = scale_param[1]), 
     type = "l", ylab = "Weibull Density")

for(k in 2:length(scale_param)){
  lines(z,dweibull(z, shape = 3, scale = scale_param[k]), col = k)
}
legend("topright", c(expression(theta ~ '= 50'), expression(theta ~ '= 100'), 
                     expression(theta ~ '= 150'), expression(theta ~ '= 200')), 
       lty = 1, col = 1:4)
title(substitute(paste("Pdf Weibull Density with", " ", alpha, " = 3", " ", 
                       "and Varying Scale")))
```


## Generalized Beta Distribution of The Second Kind (GB2) {.tabset}
This section demonstrates the effect of the shape and scale parameters on the GB2 density function.

### Varying The Scale Parameter
The graph shows the GB2 density function with varying scale parameter $(\theta)$
```{r}
# Example 4:GB2
gb2_density <- function (x, shape_1, shape_2, shape_3, scale){
  mu <- log(scale)
  sigma <- 1 / shape_3
  xt <- (log(x) - mu) / sigma
  logexpxt <- ifelse (xt > 23, yt, log(1 + exp(xt)))
  logdens <- shape_1 * xt - log(sigma) - log(beta(shape_1, shape_2)) - 
    (shape_1 + shape_2) * logexpxt - log(x) 
  exp(logdens)
}
x <- seq(0, 400, by = 1)

alpha_1 <- 5
alpha_2 <- 4 
gamma <- 2
theta <- seq(150, 250, 50)

# Varying the scale parameter
plot(x, 
     gb2_density(x, shape_1 = alpha_1, shape_2 = alpha_2, shape_3 = gamma, 
                 scale = theta[1]), type = "l", ylab = "Gen Beta 2 Density", 
     main = expression(paste("GB2 Density with ", alpha[1], " = 5, ", 
                             alpha[2], " = 4, ", alpha[3], " = 2, ", 
                             "and Varying Scale (", theta, ") Parameters")), 
     cex.main = 0.95 )

for(k in 2:length(theta)){
  lines(x, gb2_density(x, shape_1 = alpha_1, shape_2 = alpha_2, shape_3 = gamma, 
                       scale = theta[k]), col = k)
}
legend("topleft", c(expression(theta ~ '= 150'), expression(theta ~ '= 200'), 
                    expression(theta ~ '= 250')), lty = 1, cex = 0.6, col = 1:3)

```

Note: Here we wrote our own function for the density function of the GB2 density function. 

## Methods of Creating New Distributions {.tabset}
This section shows some of the methods of creating new distributions.

### Mixture Distributions
The graph below creates a density function from two random variables that follow a gamma distribution. 

```{r}
# Example 5: A mixed density
# Specify density of a mixture of 2 gamma distributions
mixture_gamma_density <- function (x, a_1, a_2, alpha_gamma1, theta_gamma1, alpha_gamma2,
                                   theta_gamma2){
  a_1 * dgamma(x, shape = alpha_gamma1, scale = theta_gamma1) + a_2 * 
    dgamma(x, shape = alpha_gamma2, scale = theta_gamma2)
}

w <- 1:30000 / 100
a_1 <- 0.5
a_2 <- 0.5
alpha_1 <- 4
theta_1 <- 7
alpha_2 <- 15
theta_2 <- 7

mix_gamma_dens <- mixture_gamma_density(w, a_1, a_2, alpha_1, theta_1, alpha_2, theta_2)

plot(w, mix_gamma_dens, type = "l")

```

### Density Obtained Through Splicing 
The graph below shows a density function through <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-trigger="hover" data-content="Join together different probability density functions to form a pdf over the support of a random variable.", style="color:green;">splicing</a> by combining an exponential distribution on $(0,c)$ with a Pareto distribution on $(c,\infty)$

```{r}
# Example 6: density obtained through splicing
# Combine an Exp on (0,c) with a Pareto on (c,\infty)

splice_exp_par <- function (x, c, v, theta, gamma, alpha){
  if (0 <= x & x < c) {return(v * dexp(x, 1 / theta) / pexp(c, 1 / theta))} else 
    if (x >= c) {return((1 - v) * dparetoII(x, loc = 0, shape = alpha, scale = theta) / 
                          (1 - pparetoII(x, loc = 0, shape = alpha, scale = theta)))}
}

x <- t(as.matrix(1:2500 / 10))

splice_values <- apply(x, 2, splice_exp_par, c = 100, v = 0.6, 
                       theta = 100, gamma = 200, alpha = 4)

plot(x, splice_values, type = 'l')
```

## Coverage Modifications 

### Load Required Package 
 The `actuar` package provides functions for dealing with coverage modifications. In the following sections we will check the functionalities of the `coverage` command. 
 
```{r warning = FALSE , message = FALSE}
library(actuar)
```

### Ordinary Deductible {.tabset}
This section plots the modified probability density functions due to deductibles for the <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-trigger="hover" data-content="Zero is observed for losses less than the deductible.", style="color:green;">payment per loss</a> and <a href="javascript://" title="" data-toggle="popover" data-placement="top" data-trigger="hover" data-content="If the loss is less than the deductible, then the insurer does not observe the loss.", style="color:green;">payment per payment</a> random variables.

#### Payment Per Loss with Ordinary Deductible
Let $X$ be the random variable for loss size. The random variable for the payment per loss with deductible $d$ is $Y^L=(X-d)_+$. The plot of the modified probability density function is below. 

```{r}
f <- coverage(dgamma, pgamma, deductible = 1, per.loss = TRUE)  # create the object
mode(f)  # it's a function. Here deductible is 1
# Check the pdf for Y^L at 0 and the original loss at 1
f(0, 3)  # mass at 0
pgamma(0 + 1, 3)  # idem

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3), lwd = 1, col = "gray")  # original
curve(dgamma(x, 3), from = 1, to = 10, ylim = c(0, 0.3), lwd = 2, add = TRUE)
curve(f(x, 3), from = 0.01, col = "blue", add = TRUE, lwd = 2)  # modified
points(0, f(0, 3), pch = 16, col = "blue")
legend("topright", c("Original pdf", "Modified pdf"), 
       lty = 1, cex = 0.6, col = c("black","blue"))

```

A few quick notes on these commands:

* The `coverage()` function computes probability density function or cumulative distribution function of the payment per payment or payment per loss random variable under any combination of the following coverage modifications: deductible, limit, coinsurance, inflation. In this illustration we used it to compute the probability density function of the payment per loss random variable with a deductible of 1. 
* The `f(0, 3)` function calculates the pdf when the payment per loss variable is 0 with gamma parameters shape=3 and rate=1. Because we used a deductible of 1 , this should be equal to `pgamma(0 + 1, 3)`. 

#### Payment Per Payment with Ordinary Deductible

$Y^P$ with pdf $f_{Y^P}(y) = f_X(y+d)/S_X(d)$

```{r}

f <- coverage(dgamma, pgamma, deductible = 1)  # create the object

f(0, 3)  # calculate in x = 0, shape = 3, rate = 1
f(5, 3)  # calculate in x = 5, shape = 3, rate = 1

dgamma(5 + 1, 3) / pgamma(1, 3, lower = FALSE)  # DIY
curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3), 
      lwd = 1, col = "gray")  # original pdf
curve(dgamma(x, 3), from = 1, to = 10, ylim = c(0, 0.3), add = TRUE, lwd = 2) 
curve(f(x, 3), from = 0.01, col = "blue", 
      add = TRUE, lwd = 2)     # modified pdf
legend("topright", c("Original pdf", "Modified pdf"), 
       lty = 1, cex = 0.6, col = c("black","blue"))

```

#### Per Payment Variable with Policy Limit, Coinsurance and Inflation 
```{r}
f <- coverage(dgamma, pgamma, deductible = 1, limit = 100, 
              coinsurance = 0.9, inflation = 0.05)  
         # create the object

f(0, 3)  # calculate in x = 0, shape = 3, rate = 1
f(5, 3)  # calculate in x = 5, shape = 3, rate = 1

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3), 
      lwd = 1, col = "gray")  # original pdf
curve(dgamma(x, 3), from = 1, to = 10, ylim = c(0, 0.3), add = TRUE, lwd = 2)
curve(f(x, 3), from = 0.01, col = "blue", add = TRUE, lwd = 2)  # modified pdf
legend("topright", c("Original pdf", "Modified pdf"), 
       lty = 1, cex = 0.6, col = c("black", "blue"))

``` 

### Franchise Deductible {.tabset}
A policy with a *franchise deductible* of $d$ pays nothing if the loss is no greater than $d$, and pays the full amount of the loss if it is greater than $d$. This section plots the pdf for the per payment and per loss random variable. 

#### Payment Per Loss with Franchise Deductible 

```{r}
# Franchise deductible 
# Per loss variable
f <- coverage(dgamma, pgamma, deductible = 1, per.loss = TRUE, franchise = TRUE)
f(0, 3)  # mass at 0
pgamma(1, 3)  # idem
f(0.5, 3)  # 0 < x < 1
f(1, 3)  # x = 1
f(5, 3)  # x > 1
dgamma(5, 3)

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3))  # original
curve(f(x, 3), from = 1.1, col = "blue", add = TRUE)  # modified
points(0, f(0, 3), pch = 16, col = "blue")  # mass at 0
curve(f(x, 3), from = 0.1, to = 1, col = "blue", add = TRUE)  # 0 < x < 1
legend("topright", c("Original pdf", "Modified pdf"), 
       lty = 1, cex = 0.6, col = c("black", "blue"))
```

Note : To use the franchise deductible , we have to add the option `franchise = TRUE` in the coverage function. 

#### Payment Per Payment with Franchise Deductible
```{r}
# Franchise deductible
# Per payment variable
f <- coverage(dgamma, pgamma, deductible = 1, franchise = TRUE)
f(0, 3)  # x = 0
f(0.5, 3)  # 0 < x < 1
f(1, 3)  # x = 1
f(5, 3)  # x > 1

dgamma(5, 3) / pgamma(1, 3, lower = FALSE)  # idem

curve(dgamma(x, 3), from = 0, to = 10, ylim = c(0, 0.3))  # original
curve(f(x, 3), from = 1.1, col = "blue", add = TRUE)  # modified
curve(f(x, 3), from = 0, to = 1, col = "blue", add = TRUE)  # 0 < x < 1
legend("topright", c("Original pdf", "Modified pdf"), 
       lty = 1, cex = 0.6, col = c("black", "blue"))
```

<script>
$(document).ready(function(){
    $('[data-toggle="popover"]').popover(); 
});
</script>

