---
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=4"
    ]
---
# Simulation 

*This file contains illustrative **R** code for computing important count distributions. When reviewing this code, you should open an **R** session, copy-and-paste the code, and see it perform. Then, you will be able to change parameters, look up commands, and so forth, as you go. *

## Simulation - Inversion Method
This section shows how to use the inversion method to simulate claims from a gamma distribution. The results below are summary statistics from the simulated data. 

```{r}
# Simulation - gamma
library(moments)
set.seed(2)  # set seed to reproduce work 
n_tot <- 20000  # number of simulations
alpha <- 2
theta <- 100
         
losses <- rgamma(n_tot, alpha, scale = theta)  
summary(losses)

k <- 0.95
percentile_loss <- quantile(losses,k)  # Kth percentile of losses 
percentile_loss

#######################################

# OR you can use this method to simulate losses 
# Fx <- runif(n_tot)
# losses <- qgamma(Fx, alpha, scale = theta)

#######################################

# For the Pareto Distribution, use
# library(VGAM)
# n_tot <- 10000  # number of simulations
# alpha <- 3
# theta <- 25000
# losses <- rparetoII(n_tot, scale = theta, shape = alpha)
# rparetoII(n_tot, scale = theta, shape = alpha) 
```
A few quick notes on these commands:

* The `rgamma()` function randomly generates data from the Gamma distribution.  In this illustration the data was generated from a gamma distribution with parameters `shape = alpha = 2` and `scale = theta = 100`. 
* The `quantile()` function provides sample quantiles corresponding to the given probabilities. Here we wanted the simulated loss data corresponding to the 95th percentile. 



## Comparing Moments from The Simulated Data to Theoretical Moments

```{r}
library(pander)
# Raw moments for k = 0.5
# Theoretical 
k <- 0.5
T_0.5 <- round(((theta^k) * gamma(alpha + k)) / gamma(alpha), 2)

# Simulated data raw moments
S_0.5 <- round(moment(losses, order = k, central = FALSE), 2)


# Raw moments for k = 1
# Theoretical 
k <- 1
T_1 <- ((theta^k) * gamma(alpha + k)) / gamma(alpha)

# Simulated data raw moments
S_1 <- round(moment(losses, order = k, central = FALSE), 2)

# Raw moments for k = 2
# Theoretical 
k <- 2
T_2 <- ((theta^k) * gamma(alpha + k)) / gamma(alpha)

#Simulated data raw moments
S_2<-round(moment(losses, order = k, central = FALSE),2)

# Raw moments for k = 3
# Theoretical 
k <- 3
T_3<-((theta^k) * gamma(alpha + k)) / gamma(alpha)

# Simulated data raw moments
S_3 <- round(moment(losses, order = k, central = FALSE), 2)

# Raw moments for k = 3

# Theoretical 
k <- 4
T_4 <- ((theta^k) * gamma(alpha + k)) / gamma(alpha)

# Simulated data raw moments
S_4 <- round(moment(losses, order = k, central = FALSE), 2)

pander(rbind(c("k", 0.5, 1, 2, 3, 4), c("Theoretical", T_0.5, T_1, T_2, T_3, T_4), 
             c("Simulated", S_0.5, S_1, S_2, S_3, S_4)))
```

A few quick notes on these commands:

* The `moment()` function computes all the sample moments of the chosen type up to a given order. In this illustration, we wanted the raw sample moments of the *kth* order.
* The `round()` function was used to round values. 

